---
title: "Challenge 3"
author: "Michelle Kairy"
format: 
  html:
    mainfont: "Helvetica"
    backgroundcolor: "#f0efed"
    fontcolor: "#624C2D"
    monobackgroundcolor: "#E9D9C3"
editor: visual
execute: 
  echo: true
  error: true
---

### Loading Libraries

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(here)
library(stringr)
```

### Loading Data

```{r}
#| message: false
hiphop <- read_csv(here::here("supporting_artifacts", "supporting_data", "hiphop.csv"))
```

### Data Cleaning

```{r}
hiphop <- hiphop |>
  mutate(sex = as.factor(sex),
         ethnic = ifelse(ethnic == "white", ethnic, "non-white"),
         subj = as.numeric(str_replace_all(subj, "p", "")),
         familiarity = as.numeric(familiarity))
```

#### 1. Comparisons

```{r}
hiphop <- hiphop |>
  distinct(subj, .keep_all = TRUE) |>
  select(sex, ethnic,  intl:unclassifiable) 
```

[When I first attempted this problem, I was so caught up in not using the same function twice (summarise) that I couldn't figure out a simple way to find the difference in means. Additionally, I wasn't sure if the greatest difference would be the relative difference or just the plain difference, which gave me different answers in part a. Looking back at this problem, I can see that calling the same function multiple times can sometimes be the most efficient way to code, and I shouldn't be afraid to do that.]{style="color: brown;"}

```{r}
hiphop_mf <- hiphop |>
  group_by(sex) |>
  summarise(across(.cols = c(intl:unclassifiable),
                   .fns = mean,
                   .names = "mean_{.col}"))
hiphop_mf |>
  summarise(across(.cols = c(mean_intl:mean_unclassifiable),
                   .fns = diff,
                   .names = "diff_{.col}"))

```

For male versus female participants, the biggest difference in number of artists known in a particular category, on average, is in vocal, where the difference between the mean known artists is 0.6882856 (males know 0.688 more vocal artists than females on average).

```{r}
hiphop_ethnic <- hiphop |>
  group_by(ethnic) |>
  summarise(across(.cols = c(intl:unclassifiable),
                   .fns = mean,
                   .names = "mean_{.col}"))

hiphop_ethnic |>
  summarise(across(.cols = c(mean_intl:mean_unclassifiable),
                   .fns = diff,
                   .names = "diff_{.col}"))
```

For white versus non-white participants, the biggest difference in number of artists known in a particular category, on average, is in the international genre, where the difference between the mean known artists is 1.592593 (non-white participants know 1.59 more international artists than white participants on average).

#### 2. Study Design Concerns

First, I am concerned about this study because the students involved come from just one university, which has a majority white student body.

Secondly, the study does not seem to have used a random sample, where the researcher simply asked students in a few classes to fill out a survey and self report demographic information.
