---
title: "Practice Activity 9.2"
author: "Michelle Kairy"
format:
  html: 
    self-contained: true
editor: visual
---

```{r}
#| message: false
#| warning: false
library(tidyverse)
```

A man named Professor Harold Hill wanders into your small town. He claims to be selling top-quality instruments for a marching band. He begins selling his instruments to all the children in town.

You are suspicious of Professor Hill, and think that perhaps he is selling instruments made of sub par materials, to scam the townsfolk.

You do some research on the weights of properly crafted brass instruments, and you learn the following facts:

-   Trombones weights are Normally distributed, with a mean weight of 4.6 pounds and a standard deviation of 0.8 pounds.

-   Trumpet weights are Uniformly distributed between 1.5 and 3.5 pounds.

-   Reed instruments (like clarinets, bassoons, and saxophones) have weights that are Chi-Square distributed with 4 degrees of freedom.

### 2. Warm- Up

```{r}
# What is the 95th percentile for trumpet weight?
qunif(0.95, min = 1.5, max = 3.5)

# What is the 10th percentile for trombone weight?
qnorm(0.10, mean = 4.6, sd = 0.8)

# About what percent of trombones do you expect to be more than 5 pounds?
pnorm(5, mean = 4.6, sd = 0.8, lower.tail = FALSE)

# About what percent of reed instruments do you expect to be more than 5 pounds?
pchisq(5, df = 4, lower.tail = FALSE)

# Simulate 100 random trombone weights. How many of them were below 4 pounds?
sum(ifelse(rnorm(100, mean = 4.6, sd = 0.8) < 4, 1, 0))
```

### 3. Catching a Con

```{r}
music_man <- function(n_tromb, n_cor, n_reed){
  
  trombones <- rnorm(n_tromb, mean = 4.6, sd = 0.8)
  cornets <- runif(n_cor, min = 1.5, max = 3.5)
  reeds <- rchisq(n_reed, df = 4)
  
  return(sum(trombones, cornets, reeds))
  
}

set.seed(1957)

my_weights <- rerun(.n = 1000, 
                    music_man(n_tromb = 76, n_cor = 110, n_reed = 542 + 318 + 175)
                    ) |> 
  flatten_dbl()
```

```{r}
sum(ifelse(my_weights < 4532, 1, 0))
```

Out of 1000 simulated total weights, only 7 had a weight less than 4532 pounds. Professor Hill is probably lying, but he seems charming, so I'll let him keep doing his thing.
